---
bibliography: [references-manuscript.bib, r-references.bib]
---

# Experiment 1: Gender categorization task

In Experiment 1, participants were shown target words flanked by articles that either agreed (were concordant with) or disagreed with the grammatical gender of the target word (discordant). They were then asked to indicate the grammatical gender of the target word. Our predictions for the outcome of this experiment were as follows: (1) We expected that this task would maximize the effect of flanker-noun gender agreement as grammatical gender was directly relevant to the task, with discordant flankers leading to longer response times and more errors than concordant flankers. (2) Additionally, there was a no-flanker control condition. We expected both response time and accuracy to be in between the concordant and discordant conditions, however, there is some evidence that the mere presence of any flanker can inhibit processing of the central word [@snell2018b]. (3) There is evidence that, in Spanish, the masculine gender has a special status as the default grammatical gender [@beatty-martínez2019; @beatty-martínez2020]. For example, most loanwords are assigned the masculine gender [e.g., *el bit vs. la web;* @cabanillas2007] and children tend to assign the masculine gender to new, ambiguous nouns [@pérez-pereira1991]. If there is a systematic difference between processing grammatically male and grammatically feminine words in general and regardless of the flanker condition, we expected to observe this in reaction time or accuracy or both. For example, masculine words may be processed faster than feminine words overall. (4) The grammatical gender of the target word may also modulate the effect of flanker-noun gender agreement. For example, it could be the case that the flankers have a stronger effect if they do not correspond to the default gender, which, in Spanish, is the masculine gender (see above). If this is the case, we might observe the flanker effect to be stronger for masculine nouns with feminine flankers than for feminine nouns with masculine flankers. Of course, the effect could also be in the opposite direction, where masculine flankers, due to masculine being the default gender, have a stronger disruptive effect on decisions on feminine words than feminine flankers on masculine words. In both of these cases, we would observe an interaction between stimulus gender and flanker condition.\

## Method 

We report here how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the experiment. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) --> The experiment was preregistered; the preregistration form can be found at <https://osf.io/5j9q2>. Data, experiment scripts, stimuli, and fully reproducible analysis code can be found at <https://github.com/bernhardangele/SerranoGomezAngele2026>. The experiment (as well as Experiments 2 and 3) was conducted in accordance with the ethical guidelines laid out in the Helsinki Declaration and was approved by the ethics boards of the California State University San Bernardino (IRB: IRB-FY2023-10) and Nebrija University (UNNE-2024-0012).

### Rationale for sample size

@brysbaert2019 established a simple rule of thumb for designs involving both random effects for participants and for language (e.g. words or sentences). They recommended a minimum of 1,600 observations per condition to reliably detect small effects. Our design has a total of six conditions (3 flanker conditions X 2 grammatical genders) and 450 items, meaning that each participant contributes 75 observations per condition. This means that 22 participants would theoretically be sufficient to reach the 1,600 observations. However, as there is no previous research on articles as flankers and we had no clear expectations regarding the size of the flanker effect (and, particularly, the interaction between the flanker manipulation and the grammatical gender of the target word), we decided to increase the target number of observations to 6,400, which would translate to a total of 86 participants. We estimated that, with a budget of \$400, we would be able to collect sufficient observations even with a moderate amount of data loss due to low accuracy and participant drop-out. We therefore set a stopping rule to finish data collection when the budget was spent. In the end, participants were more accurate and more likely to finish the experiment than expected, leaving us with more than 100 participants even after exclusions.

### Participants

```{r}
#| label: read-data-and-demographics-for-all-experiments
#| echo: false
#| message: false

exp1_all_csv_content <- fs::dir_ls(path = here("data", "exp1_gender_decision_flanker"), glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
    .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer()))

exp1_all_participants <- exp1_all_csv_content |>
  # for each subject, set the first value of counterbalancing condition as the value for all of the data
  # the key assignment is given by the image used in the instruction (3rd row of each data file) 
  group_by(PROLIFIC_PID) |>
  mutate(counterbalancing_condition = counterbalancing_condition[1],
         response_key_condition = if_else(instruct_image[3] == "keyboardresponseFM.png", "FM", "MF")) |>
  ungroup() |>
  # remove rows that are not trials (e.g., instructions) and practice trials
  filter(!is.na(TrialID) & TrialID < 2000) %>%
  # fix some misleading variable names and add response key counterbalancing information (female left vs. male left)
  dplyr::mutate(Target = Prime, 
         FlankerDuration = PrimeDuration) |>
  dplyr::select(source, PROLIFIC_PID, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, FlankerDuration, Target, Flanker, response_key_condition, counterbalancing_condition)

participant_list_exp1 <- exp1_all_participants |> group_by(PROLIFIC_PID, counterbalancing_condition, response_key_condition) |> summarise(
  mean_rt = mean(rt),
  mean_acc = mean(corr[corr != -1]),
  mean_correct = mean(corr == 1),
  time_out = sum(corr == -1),
  N = n()
  )

low_acc_participants_exp1 <- participant_list_exp1 %>% filter(mean_correct < .8)

# this is the participant who confirmed on Prolific that they were over 18 and then entered 12 as their age on the survey
# they are also a low-accuracy participant, so excluded anyway
exclude_for_inconsistent_information <- "0d828a45ab54823a65f9fca4"

exp1 <- exp1_all_participants %>% filter(!(PROLIFIC_PID %in% low_acc_participants_exp1$PROLIFIC_PID)) %>% 
  filter(!(PROLIFIC_PID %in% exclude_for_inconsistent_information))

n_exp1_correct_responses <- exp1 %>% filter(corr == 1) %>% nrow()

n_exp1_correct_responses_short <- exp1 %>% filter(corr == 1 & rt < 0.25) %>% nrow()

n_exp1_correct_responses_long <- exp1 %>% filter(corr == 1 & rt > 1.8) %>% nrow()

n_exp1_timeout_responses <- exp1 %>% 
  filter(corr == -1) %>% nrow()

n_exp1_all_responses <- exp1 %>% 
  filter(corr != -1) %>% nrow()

n_exp1_all_responses_excluding_short_and_long <- exp1 %>% 
  filter(!(rt < 0.25 | rt > 1.8) & corr != -1) %>% nrow()


exp1$Condition <- factor(exp1$Condition, levels = c("CON", "AGR", "DIS"))

exp1$StimulusType <- factor(exp1$StimulusType, levels = c("MAS", "FEM"))

demographics <- read_csv(here("analysis", "all_exp_demographics.csv"))

demographics_summary_exp1 <- demographics %>%
  filter(Experiment == "Exp1" & !(participant %in% low_acc_participants_exp1$PROLIFIC_PID)) %>%
  summarise(n_total = n(),
            mean_age = mean(age),
            median_age = median(age),
            min_age = min(age),
            max_age = max(age),
            n_male = sum(gender == "M"),
            n_female = sum(gender == "F"),
            n_nb = sum(!gender %in% c("M", "F")),
            n_first_language_spanish = sum(first_language == "Spanish"),
            n_first_language_spanish_english = sum(first_language %in% c("Spanish/English", "English/Spanish")),
            n_first_language_spanish_other = sum(first_language == "Spanish/Other")
            )


```

A total of `r nrow(participant_list_exp1) %>% apa_num()` participants were recruited through Prolific. On the Prolific platform, the experiment was only shown to participants who had indicated that they were over 18 years of age, that Spanish was their first language and that they had no diagnosis of a language-related disorder. Participants were paid \$2.50 (or the equivalent in their currency) for their participation, corresponding to \$10/hour given that the experiment lasted about 15 minutes on average. Out of all the participants, `r nrow(low_acc_participants_exp1) %>% as.integer %>% apa_num(numerals = FALSE)` were excluded from the analysis due to low accuracy (less than 80% correct responses). The final sample consisted of `r demographics_summary_exp1[1,"n_total"] %>% apa_num` participants (`r demographics_summary_exp1[1,"n_male"] %>% apa_num` male, `r demographics_summary_exp1[1,"n_female"] %>% apa_num` female, `r demographics_summary_exp1[1,"n_nb"] %>% apa_num` non-binary). The median age of the participants was `r demographics_summary_exp1[1,"median_age"] %>% apa_num(digits = 0)` years, with a minimum age of `r demographics_summary_exp1[1,"min_age"] %>% apa_num(digits = 0)` and a maximum age of `r demographics_summary_exp1[1,"max_age"] %>% apa_num(digits = 0)`. The majority of the participants stated that their first language was only Spanish (`r demographics_summary_exp1[1,"n_first_language_spanish"] %>% apa_num` participants), with `r demographics_summary_exp1[1,"n_first_language_spanish_english"] %>% apa_num` reporting both Spanish and English as their first languages and `r demographics_summary_exp1[1,"n_first_language_spanish_other"] %>% apa_num` reporting Spanish and another language as their first languages. All participants were naive to the purpose of the experiment and gave informed consent before the experiment.

### Materials

We used a total of 450 high-frequency Spanish target words (e.g., *mesa*) with four, five, and six letters. Half of the target words were feminine and the other half were masculine. @tbl-exp1-material-descriptives shows the number of target words in each category and their word frequency.

```{r}
#| label: tbl-exp1-material-descriptives
#| echo: false
#| message: false
#| tbl-cap: "Descriptive statistics of the target words used in Experiment 1."
#| after-caption-space: 0pt

# read stimuli from excel file
materials_exp1 <- readxl::read_xlsx(here("exp1_grammatical_gender_categorization_flanker", "trials_list0.xlsx"))

# write word list for EsPal if necessary
#write.csv(materials_exp1$Prime, here("exp1_grammatical_gender_categorization_flanker", "exp1_words.txt"), row.names = FALSE, quote = FALSE)

materials_exp1_freq <- read_delim(here("exp1_grammatical_gender_categorization_flanker", "espal_exp1_words.txt"))

# join frequency data with the materials

materials_exp1_full <- materials_exp1 %>%
  left_join(materials_exp1_freq, by = c("Prime" = "word")) %>%
  mutate(word = as.character(Prime),
         word_gender = if_else(StimulusType == "MAS", "Masculine", "Feminine")) %>%
  select(c(word, word_gender, Condition, frq,zipf, num_letters))

materials_exp1_descriptives_table <- materials_exp1_full %>%
  group_by(word_gender, num_letters) %>%
  summarise(mean_zipf = mean(zipf),
            sd_zipf = sd(zipf),
            N = n())

# print table as gt

materials_exp1_descriptives_table %>%
  gt(rowname_col = "num_letters") %>%
#  tab_header(
#    title = "Descriptive statistics of the target words used in Experiment 1") %>%
  cols_label(
    word_gender = "Grammatical Gender",
    num_letters = "Number of Letters",
    mean_zipf = "Mean Zipf",
    sd_zipf = "SD Zipf",
    N = "N") %>%
  fmt_number(
    columns = c(mean_zipf, sd_zipf),
    decimals = 2) %>%
  fmt_number(
    columns = c(N),
    decimals = 0
      ) %>%
  fmt_number(
    columns = c(num_letters),
    decimals = 0,
    pattern = "{x} letters"
  ) %>%
  tab_spanner(
    label = "Target Word",
    columns = c(N, mean_zipf, sd_zipf, num_letters))  %>%
  tab_source_note(
    source_note = html("Note: Word frequency is given in Zipf (\\(\\log_{10}(\\text{fpmw}) +  3 \\), where fpmw is the frequency per million words)."))

#    source_note = md("Note: Word frequency in Zipf ($ Z = \\log_{10}(\\text{fpmw}) + 3 $, #where fpmw is the frequency per million words.)"))

```

### Apparatus

Participants accessed the experiment via a link on the Prolific platform. Participants were instructed to complete the experiment on a laptop or desktop computer, and to use a physical keyboard. The experiment was presented using PsychoJS [the online Javascript-based version of PsychoPy, @peirce2019] and hosted on Pavlovia. The experiment was presented in the full screen mode of the browser, and participants were informed that the experiment would terminate if the exited full screen mode before completing the experiment.

### Procedure

The experiment was conducted online. It was presented in Spanish, and all instructions were given in Spanish. After confirming their participation on Prolific, participants were directed to the experiment. Before the start of the experiment, participants gave informed consent and confirmed that they were in a quiet environment without distractions and would be able to focus on the experiment for 15-20 minutes. They responded to a short survey with demographic data, and confirmed that they fulfilled the selection criteria (Spanish native speakers, no diagnosis of language-related disorders). The experimental procedure was very similar to that used by @snell2018b: During each experimental trial, participants first saw a fixation target (two vertical bars located above and below the center point of the screen, with the center of the screen marked by the gap between the bars) for 500 ms. The target word then appeared for 150 ms in the center of the screen, between the two vertical bars, either on its own (in the control condition) or along with the flankers to the left and right. The flankers were either articles agreeing with the noun (concordant condition, e.g., *la mesa la*) or articles that constituted an agreement violation (discordant condition, e.g., *el mesa el*). Every participant saw 225 feminine and 225 masculine target words, counterbalanced across the control, concordant, and discordant conditions. Each target word was presented the same number of times in each condition across participants. After 150 ms, both the target word and the flankers disappeared. This presentation time was short enough that participants were not able to plan and execute a saccade before the stimulus offset. Participants were instructed to make their decision solely based on the target word (ignoring the flankers) and as soon as the target word disappeared. They were asked to respond as quickly and as accurately as possible. Participants responded by using either the “A” and “L” keys on the physical keyboard of their devices. The assignment of these keys was counterbalanced such that half of the participants received one assignment (A = masculine and L = feminine) and the other half of participants received the other (A = feminine and L = masculine). After making the response, participants received feedback: a green tick mark emoji for correct responses or a red "x" emoji for incorrect responses were presented for 500 ms in the center of the screen. If no response was made within 2000 ms, the trial ended and participants received the feedback "Too slow!" ("¡Muy lento!") and the response was counted as an incorrect response. After the trial end, the experiment proceeded to the next trial until all trials had been completed. The trials were presented in random order. Participants completed 15 practice trials to familiarize themselves with the procedure and then performed the 450 experimental trials. Participants were offered the opportunity to take two breaks during the experiment, after 150 trials and after 300 trials. After the experiment, participants were thanked for their participation and returned to the Prolific platform. The experiment took about 15 minutes on average.

### Data analysis and dependent variables

We analysed the data by fitting Bayesian linear mixed models using the *brms* package [@R-brms_a; @R-brms_b; @R-brms_c] in R[^1]. These models included the flanker condition (control vs. concordant vs. discordant) as well as the grammatical gender of the target word (feminine vs. masculine) and their interaction as categorical fixed effects. We used Helmert contrasts for the flanker condition, with the first contrast comparing the discordant flanker condition (represented as -1) to the mean of the concordant and the control conditions (represented as .5 and .5) and the second contrast comparing the no-flanker control condition (represented as 1) to the concordant flanker condition (represented as -1). For the grammatical gender, we used a sum contrast comparing the feminine (-1) to the masculine grammatical gender (1). We included the maximum random effects structure [i.e., random intercepts and random slopes for condition, stimulus type, and their interaction over subjects and random intercepts and random slopes for condition over items, @barr2013]. We used the ex-Gaussian distribution to model correct response times (in s) after excluding excessively short (response time \< .25 s; `r n_exp1_correct_responses_short %>% apa_num` of `r n_exp1_correct_responses %>% apa_num` total correct responses) and long responses (response time \> 1.8 s; `r n_exp1_correct_responses_long %>% apa_num` responses), with both the mean of the Gaussian component $\mu_{RT}$ and the scale parameter of the exponential component $\beta_{RT}$ (equaling the inverse of the rate parameter $\lambda$) being allowed to vary between conditions. To model response accuracy, we used the Bernoulli distribution with a logit link after excluding no-response trials (`r n_exp1_timeout_responses %>% apa_num` trials) and short and long responses as above (resulting in `r n_exp1_all_responses_excluding_short_and_long %>% apa_num` out of `r n_exp1_all_responses %>% apa_num` no-timeout responses being included). We used the default priors suggested by *brms* except for the coefficients for the fixed effects, for which we applied weakly informative priors of $\beta \sim N(0,1)$ in order to rule out improbably large effect sizes. Each model was fitted using four chains with 5000 iterations each with 1000 warmup iterations. The models converged successfully (all $\hat{R}$s = 1.00). We report the mean, the estimates (b) and the 95% Bayesian Credible Intervals (95% CrIs) based on the posterior distribution of each model parameter. In order to simplify the interpretation of the posterior distribution, we highlight whether 0 is not a credible value for its coefficient (i.e., if it is not part of the 95% CrI). Additionally, we report the proportion of the posterior distribution on the same side of 0 (i.e., the proportion of the posterior distribution that is greater than 0 or less than 0). We always report the greater proportion (e.g., if 49% of the posterior distribution is \< 0 and 51 % is \> 0 we report .51). We also report the ratio of the two proportions (with the greater proportion always in the numerator), which effectively corresponds to a Bayes Factor for the directional hypothesis b \> 0 or b \< 0, respectively. Finally, we report the Bayes Factor (BF) calculated using the Savage-Dickey density ratio which tests the evidence for the point-null hypothesis b = 0. These measures all provide complementary information about the effect in question. We interpret BF values of 3-10 as moderate evidence, 10-30 as strong evidence, and \>30 as very strong evidence for the effect [@jeffreys1961].

[^1]: The versions of R and all packages used are as follows: `r papaja::cite_r("r-references.bib")`.

## Results

```{r}
#| label: load-exp1-model-results
#| echo: false
#| message: false
#| warning: false
#| cache: true

# load the model results
blmm_exp1_rt_dist <- qs_read(here("analysis", "blmm_exp1_rt_dist.qs"))
blmm_exp1_acc <- qs_read(here("analysis", "blmm_acc_exp1.qs"))
```

In general, and after excluding the seven low-accuracy participants (see above), accuracy was very high, with `r (n_exp1_correct_responses/n_exp1_all_responses * 100) %>% apa_num`% correct responses out of `r n_exp1_all_responses %>% apa_num` total responses. The mean response time for correct answers was `r exp1 %>% filter(corr == 1 & rt > .25 & rt < 1.8) %>% .$rt %>% mean() %>% apa_num(digits = 3)` seconds. The mean response time for incorrect answers was `r exp1 %>% filter(corr == 0 & rt > .25 & rt < 1.8) %>% .$rt %>% mean() %>% apa_num`.

@tbl-exp1-means summarizes the means and standard deviations of the correct response times by flanker condition and grammatical gender of the target word, as well as the number of observations in each condition.

```{r}
#| label: tbl-exp1-means
#| echo: false
#| message: false
#| after-caption-space: 0pt
#| tbl-cap: Accuracy as well as means and standard deviations of correct response times (in s) by flanker condition and grammatical gender of the target word in Experiment 1.

# Calculate proportion correct separately, excluding corr == -1
exp1_percent_correct <- exp1 %>%
  filter(rt > .25 & rt < 1.8) %>% # Apply the same RT filter as for means for consistency in trial count basis
  group_by(StimulusType, Condition) %>%
  summarise(
    n_trials = n(),
    n_correct = sum(corr == 1),
    percent_correct = n_correct / n_trials * 100
  ) %>%
  ungroup()

# Calculate means and standard deviations for correct RTs
exp1_means_table <- exp1 %>%
  filter(corr == 1 & rt > .25 & rt < 1.8) %>%
  group_by(StimulusType, Condition) %>%
  summarise(
    mean_rt = mean(rt),
    sd_rt = sd(rt),
    N = n() # N here is the number of correct trials within the RT filter
  ) %>%
  ungroup() %>%
  mutate(
    Condition = recode(Condition,
                       "CON" = "Control",
                       "AGR" = "Concordant",
                       "DIS" = "Discordant"),
    StimulusType = recode(StimulusType,
                          "MAS" = "Masculine",
                          "FEM" = "Feminine")
  )

# Join the proportion correct data with the means table
exp1_means_table <- exp1_means_table %>%
  left_join(exp1_percent_correct %>%
              mutate(Condition = recode(Condition,
                                        "CON" = "Control",
                                        "AGR" = "Concordant",
                                        "DIS" = "Discordant"),
                     StimulusType = recode(StimulusType,
                                           "MAS" = "Masculine",
                                           "FEM" = "Feminine")) %>%
                          select(StimulusType, Condition, percent_correct),
            by = c("StimulusType", "Condition")) %>%
  select(Condition, StimulusType, percent_correct, N, mean_rt, sd_rt)


# use gt to make a table

exp1_means_table %>%
  gt(rowname_col = "StimulusType") %>%
  cols_label(
    StimulusType = "Grammatical Gender",
    Condition = "Flanker Condition",
    N = "N",
    percent_correct = "% correct", # Keep this label, it will appear under the spanner
    mean_rt = "Mean RT",
    sd_rt = "SD"
  ) %>%
  tab_spanner( # Add spanner column for Accuracy
    label = "Accuracy",
    columns = percent_correct
  ) %>%
  tab_spanner( # Add spanner column for RT
    label = "Correct responses",
    columns = c(N, mean_rt, sd_rt)
  ) %>%
  fmt_number(
    columns = c(mean_rt, sd_rt),
    decimals = 3
  ) %>%
  fmt_number(
    columns = percent_correct,
    decimals = 1
  ) %>%
  tab_stubhead(label = "Grammatical Gender") %>%
  tab_source_note(source_note = md("Note: *N* = number of correct responses; RT = response time (in s); SD = standard deviation (across all trials)"))
```

@tbl-exp1-brms-rt-model shows the results of the Bayesian linear mixed model for response time, while @tbl-exp1-brms-acc-model shows the results of the Bayesian generalized linear mixed model for accuracy.

```{r}
#| label: tbl-exp1-brms-rt-model
#| echo: false
#| message: false
#| cache: true
#| tbl-cap: "Coefficients and 95% credible intervals from a Bayesian linear mixed model fitted to correct reaction times in Experiment 1, with flanker condition, target word grammatical gender, and their interaction as predictors for both the $\\mu$ and the $\\beta$ parameters of the ex-Gaussian distribution. Coefficients for which the 95% credible interval does not include 0 are shown in bold."
#| after-caption-space: 0pt

parameter_names <- c(
    "Intercept",
    "Intercept",
    "Control vs. Concordant Flanker",
    "Control/Concordant vs. Discordant Flanker",
    "Grammatical Gender (Feminine vs. masculine)",
    "Control vs. Concordant:Grammatical Gender",
    "Control/Concordant vs. Discordant:Grammatical Gender",
    "Control vs. Concordant Flanker",
    "Control/Concordant vs. Discordant Flanker",
    "Grammatical Gender (Feminine vs. masculine)",
    "Control vs. Concordant:Grammatical Gender",
    "Control/Concordant vs. Discordant:Grammatical Gender)"
    )

blmm_exp1_rt_dist_summary <- cbind(parameter_names, summary(blmm_exp1_rt_dist)$fixed %>% select(-Rhat, -ends_with("_ESS")))

colnames(blmm_exp1_rt_dist_summary) <- c("parameter", "estimate", "SE","lower","upper")

# makes a vector 
beta_or_not <- if_else(rownames(blmm_exp1_rt_dist_summary) %>% startsWith("beta_"), true = "beta", "mu")

gt(blmm_exp1_rt_dist_summary, rowname_col = "parameter") %>%
  tab_row_group(id = "mu", label = md("Parameter: $\\mu$"), rows = which(beta_or_not == "mu")) %>%
  tab_row_group(id = "beta", label = md("Parameter: $\\beta$"), rows = which(beta_or_not == "beta")) %>%
  row_group_order(c("mu", "beta")) %>%
  cols_label(
    parameter = "Parameter",
    estimate = "b",
    SE = "SE",
    lower = "2.5%",
    upper = "97.5%"
  ) %>%
  tab_spanner(
    label = "Estimate",
    columns = c(estimate, SE)
  ) %>%
  tab_spanner(
    label = "Credible Interval (b)",
    columns = c(lower, upper)
  ) %>%
  fmt_number(
    columns = c(estimate, SE, lower, upper),
    decimals = 3
  ) %>% tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_body(
      columns = c(estimate, SE, lower, upper),
      rows = ((lower > 0 & upper > 0) | (lower < 0 & upper < 0 ))
    )) %>%
  tab_source_note(source_note = md("Note: *b* = coefficient estimate; *SE* = standard error; *2.5%* and *97.5%* = lower and upper bounds of the 95% credible interval. Coefficients for which the 95% credible interval does not include 0 are shown in **bold**."))
      

```

### Effects of the flanker condition

```{r}
#| label: calculate-exp1-effects
#| cache: true
#| include: false

# pre-generate all strings we need to report the experiment so we can cache this operation 


# --- Experiment 1: Response Time Model (blmm_exp1_rt_dist) ---

# Mu parameters (Location)
exp1_eff_rt_mu_con_vs_agr <- blmm_exp1_rt_dist %>% brms_effect_summary("ConditionCON_vs_AGR")
exp1_eff_rt_mu_con_agr_vs_dis <- blmm_exp1_rt_dist %>% brms_effect_summary("ConditionCON_AGR_vs_DIS")
exp1_eff_rt_mu_stimulus <- blmm_exp1_rt_dist %>% brms_effect_summary("StimulusType1")
exp1_eff_rt_mu_con_vs_agr_by_stimulus <- blmm_exp1_rt_dist %>% brms_effect_summary("ConditionCON_vs_AGR:StimulusType1")
exp1_eff_rt_mu_con_agr_vs_dis_by_stimulus <- blmm_exp1_rt_dist %>% brms_effect_summary("ConditionCON_AGR_vs_DIS:StimulusType1")

# Beta parameters (Scale/Tail)
exp1_eff_rt_beta_con_vs_agr <- blmm_exp1_rt_dist %>% brms_effect_summary("beta_ConditionCON_vs_AGR")
exp1_eff_rt_beta_con_agr_vs_dis <- blmm_exp1_rt_dist %>% brms_effect_summary("beta_ConditionCON_AGR_vs_DIS")
exp1_eff_rt_beta_stimulus <- blmm_exp1_rt_dist %>% brms_effect_summary("beta_StimulusType1")
exp1_eff_rt_beta_con_vs_agr_by_stimulus <- blmm_exp1_rt_dist %>% brms_effect_summary("beta_ConditionCON_vs_AGR:StimulusType1")
exp1_eff_rt_beta_con_agr_vs_dis_by_stimulus <- blmm_exp1_rt_dist %>% brms_effect_summary("beta_ConditionCON_AGR_vs_DIS:StimulusType1")


# --- Experiment 1: Accuracy Model (blmm_exp1_acc) ---

exp1_eff_acc_con_vs_agr <- blmm_exp1_acc %>% brms_effect_summary("ConditionCON_vs_AGR")
exp1_eff_acc_con_agr_vs_dis <- blmm_exp1_acc %>% brms_effect_summary("ConditionCON_AGR_vs_DIS")
exp1_eff_acc_stimulus <- blmm_exp1_acc %>% brms_effect_summary("StimulusType1")
exp1_eff_acc_con_vs_agr_by_stimulus <- blmm_exp1_acc %>% brms_effect_summary("ConditionCON_vs_AGR:StimulusType1")
exp1_eff_acc_con_agr_vs_dis_by_stimulus <- blmm_exp1_acc %>% brms_effect_summary("ConditionCON_AGR_vs_DIS:StimulusType1")
```

In response time, we observed a substantial effect of the discordant flanker condition compared to the two other conditions in both the mean of the Gaussian component, $\mu_{RT}$ (`r exp1_eff_rt_mu_con_agr_vs_dis`), indicating longer response times in the discordant condition, and the scale component, $\beta_{RT}$ (`r exp1_eff_rt_beta_con_agr_vs_dis`), indicating a longer right tail of the response time distribution. Participants also made more errors in the discordant compared to the concordant condition (`r exp1_eff_acc_con_agr_vs_dis`). The effect of the concordant flanker condition compared to the control condition was much smaller in $\mu_{RT}$ (`r exp1_eff_rt_mu_con_vs_agr`) and virtually absent in $\beta_{RT}$ (`r exp1_eff_rt_beta_con_vs_agr`), indicating that the concordant condition was not much different from the control condition in terms of response time. However, participants tended to be slightly more accurate in the concordant condition compared to the control condition (`r exp1_eff_acc_con_vs_agr`).

```{r}
#| label: tbl-exp1-brms-acc-model
#| echo: false
#| message: false
#| cache: true
#| tbl-cap: "Coefficients and 95% credible intervals from a Bayesian generalized linear mixed model using the Bernoulli distribution to predict response accuracy in Experiment 1, with target word grammatical gender, and their interaction as predictors. Coefficients for which the 95% credible interval does not include 0 are shown in bold."
#| after-caption-space: 0pt

parameter_names <- c(
    "Intercept",
    "Control vs. Concordant Flanker",
    "Control/Concordant vs. Discordant Flanker",
    "Grammatical Gender (Feminine vs. masculine)",
    "Control vs. Concordant:Grammatical Gender",
    "Control/Concordant vs. Discordant:Grammatical Gender"
    )

blmm_exp1_acc_summary <- cbind(parameter_names, summary(blmm_exp1_acc)$fixed %>% select(-Rhat, -ends_with("_ESS"))) 

colnames(blmm_exp1_acc_summary) <- c("parameter", "estimate", "SE","lower","upper")

gt(blmm_exp1_acc_summary, rowname_col = "parameter") %>%
  tab_row_group(id = "theta", label = md("Parameter: $\\theta$"), rows = everything()) %>%
  row_group_order(c("theta")) %>%
  cols_label(
    parameter = "Parameter",
    estimate = "b",
    SE = "SE",
    lower = "2.5%",
    upper = "97.5%"
  ) %>%
  tab_spanner(
    label = "Estimate",
    columns = c(estimate, SE)
  ) %>%
  tab_spanner(
    label = "Credible Interval (b)",
    columns = c(lower, upper)
  ) %>%
  fmt_number(
    columns = c(estimate, SE, lower, upper),
    decimals = 3
  ) %>% tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_body(
      columns = c(estimate, SE, lower, upper),
      rows = ((lower > 0 & upper > 0) | (lower < 0 & upper < 0 ))
    )) %>%
  tab_source_note(source_note = md("Note: $\\theta = \\log\\left(\\frac{p}{1-p}\\right)$, where *p* is the probability of a correct response; *b* = coefficient estimate, *SE* = standard error, *2.5%* and *97.5%* = lower and upper bounds of the 95% credible interval. Coefficients for which the 95% credible interval does not include 0 are shown in **bold**."))
      

```

### Effects of grammatical gender

We found small marginal effects of grammatical gender across flanker conditions: Response times were slightly lower for feminine target words compared to masculine target words ($\mu_{RT}$: `r exp1_eff_rt_mu_stimulus`; $\beta_{RT}$: `r exp1_eff_rt_beta_stimulus`), but accuracy was slightly higher for masculine compared to feminine target words (`r exp1_eff_acc_stimulus`).

### Interaction between flanker condition and grammatical gender

We did not observe meaningful interactions between grammatical gender of the target word and the flanker condition in response time, but, in terms of accuracy, there was an interaction between grammatical gender of the target word and the contrast comparing the concordant and control flanker conditions, with there being a larger increase in accuracy in the concordant condition compared to the control condition for feminine target words compared to masculine target words (`r exp1_eff_acc_con_vs_agr_by_stimulus`). This interaction has to be interpreted with caution as, even though 98% of the posterior distribution is above 0, the $BF_{10}$ indicates weak evidence in favor of the null hypothesis.

### Post-hoc analysis: Grammatical gender transparency

In this experiment, some of the target words were transparent in terms of gender (feminine nouns ending in -a and masculine nouns ending in “-o”, 345 target words) and some target words were opaque (nouns ending in other letters, 105 target words). We performed an additional, not pre-registered analysis involving only the transparent target words and did not find a difference in the pattern of results compared to the full dataset.

## Discussion

In the grammatical gender categorization task, we found clear evidence for an effect of the article flankers. While the concordant flankers did not affect reaction times much compared to the control condition, and only led to a small increase in accuracy compared to the control condition, the discordant flankers had a major disruptive effect on processing the target word's grammatical gender. This suggests that the article flankers are obligatorily processed and that the conflict between the discordant flanker and the correct article has a major effect on performance on the task. This is in line with the original interpretation of the flanker effect by @eriksen1974 as an inhibition of response selection by unrelated noise. It is also in line with the model proposed by @sá-leite2024. We found some evidence in favor of a bias in gender assignment across flanker conditions, but it went in the opposite direction compared to the default gender effect suggested by @beatty-martínez2019, with feminine target words being processed slightly faster but slightly less accurately than masculine target words, especially in the no-flanker control condition. Overall, however, these biases were very small compared to the effect of the discordant flankers. In summary, Experiment 1 shows clear parafoveal processing of article flankers during a grammatical gender categorization task, with participants being unable to ignore the parafoveal article information even when it disrupts the task. This is compatible with the assumption that articles and nouns are processed jointly and with the model by @sá-leite2024.
