---
bibliography: [references-manuscript.bib, r-references.bib]
---
  
  
# Experiment 3

In Experiment 3, participants were shown target words flanked by articles that either agreed or disagreed with the grammatical gender of the target word and were asked to indicate whether the target words referred to animals or not. After the grammatical categorization task in Experiment 1, which required participants to make an explicit response about the grammatical gender of the target word, and Experiment 2, which required participants to decide whether the letter string flanked by articles was a word or not, we hypothesized that this final task would require the least explicit or implicit processing of grammatical gender out of our three tasks. However, as described above, many models of word recognition (such as the Interactive Activation Model) assume that a word needs to be identified (i.e., lexical access needs to occur) before semantic information about it can be retrieved. Experiment 2 suggests that the discordant flankers can interfere with lexical access. If they also interfere with the retrieval of semantic properties of a word, we would (1) expect discordant flankers leading to longer response times and more errors than concordant flankers on the semantic categorization task. (2) There may also be a systematic difference between processing grammatically masculine and grammatically feminine words syntactically, which we may observe in response time or accuracy or both. (3) If the grammatical gender of the target word influences how participants process the flankers, and if this affects semantic processing of the target word, we expected that we would observe an interaction between word gender and flanker effect. 

## Method

### Rationale for sample size

In Experiment 3, we showed each participant a total of 416 non-animal target nouns, 208 of them masculine and 208 feminine  (we did not plan on analyzing the 140 animal noun trials). As in Experiment 2, there were two flanker conditions (concordant vs. discordant), resulting in a total of 104 observations per participant and condition. Using the rule of thumb established by @brysbaert2019, this means that 16 participants would theoretically be sufficient to reach the 1,600 observations. Given the smaller effect observed in Experiment 2 compared to Experiment 1, we suspected that the effect size in Experiment 3 could be even smaller. Because of this,  we set our target number of observations per condition (flanker condition * word gender) to 6,400. This corresponds to 60 participants. Since there was likely to be some data loss/drop out, we collected data from slightly more than 60 participants to account for potential data loss and exclusions, e.g. due to low response accuracy and to ensure a balanced design.
We report here how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the two experiments. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) --> Like Experiments 1 and 2, this experiment was preregistered; the preregistration form can be found at <https://osf.io/ektx5?view_only=abfce462e191492da94bfacf1d40fc96>. Data, experiment scripts, and stimuli can be found at <https://osf.io/hvds9/?view_only=ee68fa24f4544598bcfaffb415674ca6>. This link also contains the final analysis scripts and the Quarto code for this manuscript.


### Participants

```{r}
#| label: read-data-and-demographics-for-exp3
#| echo: false
#| message: false

exp3_all_csv_content <- fs::dir_ls(path = here("data", "exp3_semantic_categorization_flanker"), glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
    .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer()))

exp3_all_participants <- exp3_all_csv_content |>
  # for each subject, set the first value of counterbalancing condition as the value for all of the data
  # the key assignment is given by the image used in the instruction (3rd row of each data file) 
  group_by(PROLIFIC_PID) |>
  mutate(counterbalancing_condition = counterbalancing_condition[1],
         response_key_condition = if_else(instruct_image[3] == "keyboardresponseAN.png", "AN", "NA")) |>
  ungroup() |>
  # remove rows that are not trials (e.g., instructions) and practice trials
  filter(!is.na(TrialID) & TrialID < 2000) %>%
  # fix some misleading variable names and add response key counterbalancing information (female left vs. male left)
  mutate(Target = Prime, 
         FlankerDuration = PrimeDuration) |>
  select(source, PROLIFIC_PID, date, OS, frameRate, rt, corr, TrialID, StimulusType, IsAnimal, Condition, FlankerDuration, Target, Flanker, response_key_condition, counterbalancing_condition)

participant_list_exp3 <- exp3_all_participants |> group_by(PROLIFIC_PID, counterbalancing_condition, response_key_condition) |> summarise(
  mean_rt = mean(rt, na.rm = TRUE),
  mean_acc = mean(corr[corr != -1]),
  mean_correct = mean(corr == 1),
  time_out = sum(corr == -1),
  N = n()
  )

low_acc_participants_exp3 <- participant_list_exp3 %>% filter(mean_correct < .8)



exp3 <- exp3_all_participants %>% filter(!(PROLIFIC_PID %in% low_acc_participants_exp3$PROLIFIC_PID)) %>% 
  filter(!(PROLIFIC_PID %in% exclude_for_inconsistent_information))

n_exp3_correct_responses <- exp3 %>% filter(corr == 1) %>% nrow()

n_exp3_correct_responses_short <- exp3 %>% filter(corr == 1 & rt < 0.25) %>% nrow()

n_exp3_correct_responses_long <- exp3 %>% filter(corr == 1 & rt > 1.8) %>% nrow()

n_exp3_timeout_responses <- exp3 %>% 
  filter(corr == -1) %>% nrow()

n_exp3_all_responses <- exp3 %>% 
  filter(corr != -1) %>% nrow()

n_exp3_all_responses_excluding_short_and_long <- exp3 %>% 
  filter(!(rt < 0.25 | rt > 1.8) & corr != -1) %>% nrow()


exp3$Condition <- factor(exp3$Condition, levels = c("AGR", "DIS"))

exp3$Gender <- factor(exp3$StimulusType, levels = c("MAS", "FEM"))

exp3$IsAnimal <- factor(exp3$IsAnimal, levels = c("0", "1"))

demographics <- read_csv(here("analysis", "all_exp_demographics.csv"))

demographics_summary_exp3 <- demographics %>%
  filter(Experiment == "Exp3" & !(participant %in% low_acc_participants_exp3$PROLIFIC_PID)) %>%
  summarise(n_total = n(),
            mean_age = mean(age),
            median_age = median(age),
            min_age = min(age),
            max_age = max(age),
            n_male = sum(gender == "M"),
            n_female = sum(gender == "F"),
            n_nb = sum(!gender %in% c("M", "F")),
            n_first_language_spanish = sum(first_language == "Spanish"),
            n_first_language_spanish_english = sum(first_language %in% c("Spanish/English", "English/Spanish")),
            n_first_language_spanish_other = sum(first_language == "Spanish/Other")
            )


```

A total of `r nrow(participant_list_exp3) %>% apa_num()` participants were recruited through Prolific. On the Prolific platform, the experiment was only shown to participants who had indicated that they were over 18 years of age, that Spanish was their first language and that they had no diagnosis of a language-related disorder. Participants were paid \$3 (or the equivalent in their currency) for their participation, corresponding to \$9/hour given that the experiment lasted about 20 minutes on average. Out of all the participants, `r nrow(low_acc_participants_exp3) %>% as.integer %>% apa_num(numerals = FALSE)` were excluded from the analysis due to low accuracy (less than 80% correct responses). The final sample consisted of `r demographics_summary_exp3[1,"n_total"] %>% apa_num` participants (`r demographics_summary_exp3[1,"n_male"] %>% apa_num` male, `r demographics_summary_exp3[1,"n_female"] %>% apa_num` female, `r demographics_summary_exp3[1,"n_nb"] %>% apa_num` non-binary). The median age of the participants was `r demographics_summary_exp3[1,"median_age"] %>% apa_num(digits = 0)` years, with a minimum age of `r demographics_summary_exp3[1,"min_age"] %>% apa_num(digits = 0)` and a maximum age of `r demographics_summary_exp3[1,"max_age"] %>% apa_num(digits = 0)`. The majority of the participants stated that their first language was only Spanish (`r demographics_summary_exp3[1,"n_first_language_spanish"] %>% apa_num` participants), with `r demographics_summary_exp3[1,"n_first_language_spanish_english"] %>% apa_num` reporting both Spanish and English as their first languages and `r demographics_summary_exp3[1,"n_first_language_spanish_other"] %>% apa_num` reporting Spanish and another language as their first languages. As in Experiments 1 and 2, all participants were naive to the purpose of the experiment and gave informed consent before the experiment.

### Materials

We used the same stimuli as in Experiment 1, with some ambiguous words that could refer refer to both animals and non-animals (e.g., "manta") removed and more animal words added, for a total of 416 non-animal (208 feminine and 208 masculine) and 140 animal trials (70 feminine and 70 masculine). No non-animal word was repeated, but the 70 unique animal words were each presented twice to ensure there were sufficient trials (25.18%) that required an "animal" response. @tbl-exp3-material-descriptives shows the number of target words in each category and their word frequency as well as the OLD20 [@yarkoni2008]. 

```{r}
#| label: tbl-exp3-material-descriptives
#| echo: false
#| message: false
#| tbl-cap: "Descriptive statistics of the target (non-animal) words used in Experiment 3."
#| after-caption-space: 0pt

# read stimuli from excel file
materials_exp3 <- readxl::read_xlsx(here("exp3_semantic_flanker", "trials_list0.xlsx"))

#materials_exp3_not_animal <- filter(materials_exp3, IsAnimal == 0)

# write word list for EsPal if necessary
# write.csv(materials_exp3_words$Prime , here("exp3_ldt_flanker", "exp3_words.txt"), row.names = FALSE, quote = FALSE)

materials_exp3_freq <- read_delim(here("exp3_semantic_flanker", "espal_results_exp3.txt"), delim = "\t")
# remove duplicate rows
materials_exp3_freq <- materials_exp3_freq %>%
  distinct(word, .keep_all = TRUE) %>%
  select(word, frq, zipf, Lev_N, num_letters)

# join frequency data with the materials

materials_exp3_full <- materials_exp3 %>%
  left_join(materials_exp3_freq, by = c("Prime" = "word")) %>%
  mutate(word = as.character(Prime),
         word_gender = if_else(StimulusType == "MAS", "Masculine", "Feminine")) %>%
  select(c(word, word_gender, IsAnimal, Condition, frq,zipf, Lev_N, num_letters))

materials_exp3_descriptives_table <- materials_exp3_full %>%
  filter(!IsAnimal) %>% 
  group_by(word_gender, num_letters) %>%
  summarise(mean_zipf = mean(zipf),
            sd_zipf = sd(zipf),
            mean_old20 = mean(Lev_N),
            sd_old20 = sd(Lev_N),
            N = n())

# print table as gt

materials_exp3_descriptives_table %>%
  gt(rowname_col = "num_letters", groupname_col = "word_gender") %>%
#  tab_header(
#    title = "Descriptive statistics of the target words used in Experiment 1") %>%
  cols_label(
    word_gender = "Grammatical Gender",
    num_letters = "Number of Letters",
    mean_zipf = "Mean",
    sd_zipf = "SD",
    mean_old20 = "Mean",
    sd_old20 = "SD",
    N = "N") %>%
  fmt_number(
    columns = c(mean_zipf, sd_zipf, mean_old20, sd_old20),
    decimals = 2) %>%
  fmt_number(
    columns = c(N),
    decimals = 0
      ) %>%
  fmt_number(
    columns = c(num_letters),
    decimals = 0,
    pattern = "{x} letters"
  ) %>%
  tab_spanner(
    label = "Zipf",
    columns = c(mean_zipf, sd_zipf))  %>%
  tab_spanner(
    label = "OLD20",
    columns = c(mean_old20, sd_old20))  %>%
    tab_spanner(
    label = "Target Word",
    columns = c(N, mean_zipf, sd_zipf, mean_old20, sd_old20, num_letters))  %>%
  tab_source_note(
    source_note = html("Note: Word frequency is given in Zipf (\\(\\log_{10}(\\text{fpmw}) +  3 \\), where fpmw is the frequency per million words). OLD20 is the mean Levenshtein distance between the target word and the 20 nearest neighbors in the lexicon.")) 

#    source_note = md("Note: Word frequency in Zipf ($ Z = \\log_{10}(\\text{fpmw}) + 3 $, #where fpmw is the frequency per million words.)"))

```

### Apparatus

The apparatus was identical to that used in Experiments 1 and 2.

### Procedure

The procedure was identical to that used in Experiment 1 and 2, except that participants performed a semantic categorization instead of a grammatical gender categorization or lexical decision. They responded by using either the “S” and “L” keys on the physical keyboard of their devices (we avoided the "A" key in Experiment 3 due to the potential association between "A" and "animal"). As in Experiment 1, the assignment of these keys was counterbalanced such that half of the participants received one assignment (S = animal and L = non-animal) and the other half of participants received the other (S = non-animal and L = animal). As in Experiment 2, we only included the concordant and discordant flanker conditions, not the non-flanker control.

### Data analysis and dependent variables

As in Experiments 1 and 2, we analyzed the data by fitting Bayesian linear mixed models including the flanker condition (concordant vs. discordant) as well as the grammatical gender of the target word (feminine vs. masculine) and their interaction as categorical fixed effects. We only included non-animal trials in the analysis, as "yes" responses may carry an extra cost. As in Experiment 2, we used a sum contrast for the flanker condition, with the discordant flanker condition represented as -1 and the concordant flanker condition represented as 1. For the grammatical gender, we again used a sum contrast comparing the feminine (-1) to the masculine grammatical gender (1). The rest of the model specifications were identical to those used in Experiment, including the maximal random effects structure. All models converged successfully (all $\hat{R}$s = 1.00). We report the results as in Experiments 1 and 2.


## Results

```{r}
#| label: load-exp3-model-results
#| echo: false
#| message: false
#| warning: false
#| cache: true

# load the model results
blmm_exp3_rt_dist <- qs_read(here("analysis", "blmm_exp3_rt_dist.qs"))
blmm_exp3_acc <- qs_read(here("analysis", "blmm_acc_exp3.qs"))
```

As in Experiments 1 and 2, accuracy after excluding the low-accuracy participants (see above) was very high, with  `r (n_exp3_correct_responses/n_exp3_all_responses * 100) %>% apa_num`%  correct responses out of `r n_exp3_all_responses %>% apa_num` total responses (for an average of `r (n_exp3_correct_responses/8) %>% apa_num` included observations per condition). The mean response time for correct answers was `r exp2 %>% filter(corr == 1 & rt > .25 & rt < 1.8) %>% .$rt %>% mean() %>% apa_num` seconds. The mean response time for incorrect answers was `r exp2 %>% filter(corr == 0 & rt > .25 & rt < 1.8) %>% .$rt %>% mean() %>% apa_num`.

@tbl-exp3-means-non-animals summarizes the means and standard deviations of the correct response times by flanker condition and grammatical gender of target words, as well as the number of observations in each condition.

```{r}
#| label: tbl-exp3-means-non-animals
#| echo: false
#| message: false
#| after-caption-space: 0pt
#| tbl-cap: Accuracy as well as means and standard deviations of correct response times (in s) by flanker condition and grammatical gender of non-animal target words in Experiment 3.


# Calculate proportion correct separately, excluding corr == -1
exp3_percent_correct_non_animals <- exp3 %>% filter(IsAnimal == "0") %>%
  filter(rt > .25 & rt < 1.8) %>% # Apply the same RT filter as for means for consistency in trial count basis
  group_by(Gender, Condition) %>%
  summarise(
    n_trials = n(),
    n_correct = sum(corr == 1),
    percent_correct = n_correct / n_trials * 100
  ) %>%
  ungroup()

# Calculate means and standard deviations for correct RTs
exp3_means_table_non_animals <- exp3 %>%  filter(IsAnimal == "0") %>%
  filter(corr == 1 & rt > .25 & rt < 1.8) %>%
  group_by(Gender, Condition) %>%
  summarise(
    mean_rt = mean(rt),
    sd_rt = sd(rt),
    N = n() # N here is the number of correct trials within the RT filter
  ) %>%
  ungroup() 

# Join the proportion correct data with the means table
exp3_means_table_non_animals <- exp3_means_table_non_animals %>%
  left_join(exp3_percent_correct_non_animals, by = c("Gender", "Condition")) %>%
   mutate(Condition = recode(Condition,
                             "AGR" = "Concordant",
                             "DIS" = "Discordant"),
                     Gender = recode(Gender,
                                     "MAS" = "Masculine",
                                     "FEM" = "Feminine")) %>%
  select(Condition, Gender, percent_correct, N, mean_rt, sd_rt)


# use gt to make a table

gt_exp3_means_table_non_animals <- exp3_means_table_non_animals %>%
  gt(rowname_col = "Gender") %>%
  cols_label(
    Gender = "Grammatical Gender",
    Condition = "Flanker Condition",
    N = "N",
    percent_correct = "% correct", # Keep this label, it will appear under the spanner
    mean_rt = "Mean RT",
    sd_rt = "SD"
  ) %>%
  tab_spanner( # Add spanner column for Accuracy
    label = "Accuracy",
    columns = percent_correct
  ) %>%
  tab_spanner( # Add spanner column for RT
    label = "Correct responses",
    columns = c(N, mean_rt, sd_rt)
  ) %>%
  fmt_number(
    columns = c(mean_rt, sd_rt),
    decimals = 3
  ) %>%
  fmt_number(
    columns = percent_correct,
    decimals = 1
  ) %>%
  tab_stubhead(label = "Grammatical Gender") %>%
  tab_source_note(source_note = md("Note: *N* = number of correct responses; RT = response time (in s); SD = standard deviation (across all trials)"))

# Calculate proportion correct separately, excluding corr == -1
exp3_percent_correct_animals <- exp3 %>% filter(IsAnimal == "1") %>%
  filter(rt > .25 & rt < 1.8) %>% # Apply the same RT filter as for means for consistency in trial count basis
  group_by(Gender, Condition) %>%
  summarise(
    n_trials = n(),
    n_correct = sum(corr == 1),
    percent_correct = n_correct / n_trials * 100
  ) %>%
  ungroup()

# Calculate means and standard deviations for correct RTs
exp3_means_table_animals <- exp3 %>%  filter(IsAnimal == "1") %>%
  filter(corr == 1 & rt > .25 & rt < 1.8) %>%
  group_by(Gender, Condition) %>%
  summarise(
    mean_rt = mean(rt),
    sd_rt = sd(rt),
    N = n() # N here is the number of correct trials within the RT filter
  ) %>%
  ungroup() 

# Join the proportion correct data with the means table
exp3_means_table_animals <- exp3_means_table_animals %>%
  left_join(exp3_percent_correct_animals, by = c("Gender", "Condition")) %>%
   mutate(Condition = recode(Condition,
                             "CON" = "Control",
                             "AGR" = "Concordant",
                             "DIS" = "Discordant"),
                     Gender = recode(Gender,
                                     "MAS" = "Masculine",
                                     "FEM" = "Feminine")) %>%
  select(Condition, Gender, percent_correct, N, mean_rt, sd_rt)


# use gt to make a table

gt_exp3_means_table_animals <- exp3_means_table_animals %>%
  gt(rowname_col = "Gender") %>%
  cols_label(
    Gender = "Grammatical Gender",
    Condition = "Flanker Condition",
    N = "N",
    percent_correct = "% correct", # Keep this label, it will appear under the spanner
    mean_rt = "Mean RT",
    sd_rt = "SD"
  ) %>%
  tab_spanner( # Add spanner column for Accuracy
    label = "Accuracy",
    columns = percent_correct
  ) %>%
  tab_spanner( # Add spanner column for RT
    label = "Correct responses",
    columns = c(N, mean_rt, sd_rt)
  ) %>%
  fmt_number(
    columns = c(mean_rt, sd_rt),
    decimals = 3
  ) %>%
  fmt_number(
    columns = percent_correct,
    decimals = 1
  ) %>%
  tab_stubhead(label = "Apparent Grammatical Gender") %>%
  tab_source_note(source_note = md("Note: *N* = number of correct responses; RT = response time (in s); SD = standard deviation (across all trials)"))

gt_exp3_means_table_non_animals

```

@tbl-exp3-brms-rt-model-non-animals shows the results of the Bayesian linear mixed model for response time for non-animal target words, while @tbl-exp3-brms-acc-model-non-animals shows the results of the Bayesian generalized linear mixed model for accuracy (i.e. the proportion of correct rejections of target words as non-animals).

```{r}
#| label: tbl-exp3-brms-rt-model-non-animals
#| echo: false
#| message: false
#| cache: true
#| tbl-cap: "Coefficients and 95% credible intervals from a Bayesian linear mixed model fitted to correct reaction times on non-animal trials in Experiment 3, with flanker condition, target word grammatical gender, and their interaction as predictors for both the $\\mu$ and the $\\beta$ parameters of the ex-Gaussian distribution. Coefficients for which the 95% credible interval does not include 0 are shown in bold."
#| after-caption-space: 0pt

parameter_names <- c(
    "Intercept",
    "Intercept",
    "Concordant vs. Discordant Flanker",
    "Grammatical Gender (Feminine vs. masculine)",
    "Concordant vs. Discordant:Grammatical Gender",
    "Concordant vs. Discordant Flanker",
    "Grammatical Gender (Feminine vs. masculine)",
    "Concordant vs. Discordant:Grammatical Gender)"
    )

blmm_exp3_rt_dist_summary <- cbind(parameter_names, summary(blmm_exp3_rt_dist)$fixed %>% select(-Rhat, -ends_with("_ESS")))

colnames(blmm_exp3_rt_dist_summary) <- c("parameter", "estimate", "SE","lower","upper")

# makes a vector 
beta_or_not <- if_else(rownames(blmm_exp3_rt_dist_summary) %>% startsWith("beta_"), true = "beta", "mu")

gt(blmm_exp3_rt_dist_summary, rowname_col = "parameter") %>%
  tab_row_group(id = "mu", label = md("Parameter: $\\mu$"), rows = which(beta_or_not == "mu")) %>%
  tab_row_group(id = "beta", label = md("Parameter: $\\beta$"), rows = which(beta_or_not == "beta")) %>%
  row_group_order(c("mu", "beta")) %>%
  cols_label(
    parameter = "Parameter",
    estimate = "b",
    SE = "SE",
    lower = "2.5%",
    upper = "97.5%"
  ) %>%
  tab_spanner(
    label = "Estimate",
    columns = c(estimate, SE)
  ) %>%
  tab_spanner(
    label = "Credible Interval (b)",
    columns = c(lower, upper)
  ) %>%
  fmt_number(
    columns = c(estimate, SE, lower, upper),
    decimals = 3
  ) %>% tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_body(
      columns = c(estimate, SE, lower, upper),
      rows = ((lower > 0 & upper > 0) | (lower < 0 & upper < 0 ))
    )) %>%
  tab_source_note(source_note = md("Note: *b* = coefficient estimate; *SE* = standard error; *2.5%* and *97.5%* = lower and upper bounds of the 95% credible interval. Coefficients for which the 95% credible interval does not include 0 are shown in **bold**."))
      

```


```{r}
#| label: tbl-exp3-brms-acc-model-non-animals
#| echo: false
#| message: false
#| cache: true
#| tbl-cap: "Coefficients and 95% credible intervals from a Bayesian generalized linear mixed model using the Bernoulli distribution to predict response accuracy on non-animal trials in Experiment 3, with flanker condition, target word grammatical gender, and their interaction as predictors. Coefficients for which the 95% credible interval does not include 0 are shown in bold."
#| after-caption-space: 0pt


parameter_names <- c(
    "Intercept",
    "Concordant vs. Discordant Flanker",
    "Grammatical Gender (F vs. M)",
    "Concordant vs. Discordant:Grammatical Gender"
    )

blmm_exp2_acc_nonwords_summary <- cbind(parameter_names, summary(blmm_exp2_acc_nonwords)$fixed %>% select(-Rhat, -ends_with("_ESS"))) 

colnames(blmm_exp2_acc_nonwords_summary) <- c("parameter", "estimate", "SE","lower","upper")

gt(blmm_exp2_acc_nonwords_summary, rowname_col = "parameter") %>%
  tab_row_group(id = "theta", label = md("Parameter: $\\theta$"), rows = everything()) %>%
  row_group_order(c("theta")) %>%
  cols_label(
    parameter = "Parameter",
    estimate = "b",
    SE = "SE",
    lower = "2.5%",
    upper = "97.5%"
  ) %>%
  tab_spanner(
    label = "Estimate",
    columns = c(estimate, SE)
  ) %>%
  tab_spanner(
    label = "Credible Interval (b)",
    columns = c(lower, upper)
  ) %>%
  fmt_number(
    columns = c(estimate, SE, lower, upper),
    decimals = 3
  ) %>% tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_body(
      columns = c(estimate, SE, lower, upper),
      rows = ((lower > 0 & upper > 0) | (lower < 0 & upper < 0 ))
    )) %>%
  tab_source_note(source_note = md("Note: $\\theta = \\log\\left(\\frac{p}{1-p}\\right)$, where *p* is the probability of a correct response; *b* = coefficient estimate, *SE* = standard error, *2.5%* and *97.5%* = lower and upper bounds of the 95% credible interval. Coefficients for which the 95% credible interval does not include 0 are shown in **bold**."))

```


### Effects of the flanker condition

```{r}
#| label: calculate-exp3-effects
#| cache: true
#| include: false

# pre-generate all strings we need to report the experiment so we can cache this operation 


# --- Experiment 3: Response Time Model for Words (blmm_exp3_rt_dist) ---

# Mu parameters (Location)
exp3_eff_rt_mu_agr_vs_dis <- blmm_exp3_rt_dist %>% brms_effect_summary("Condition1")
exp3_eff_rt_mu_gender <- blmm_exp3_rt_dist %>% brms_effect_summary("StimulusType1")
exp3_eff_rt_mu_agr_vs_dis_by_gender <- blmm_exp3_rt_dist %>% brms_effect_summary("Condition1:StimulusType1")


# Beta parameters (Scale/Tail)
exp3_eff_rt_beta_agr_vs_dis <- blmm_exp3_rt_dist %>% brms_effect_summary("beta_Condition1")
exp3_eff_rt_beta_gender <- blmm_exp3_rt_dist %>% brms_effect_summary("beta_StimulusType1")
exp3_eff_rt_beta_agr_vs_dis_by_gender <- blmm_exp3_rt_dist %>% brms_effect_summary("beta_Condition1:StimulusType1")


# --- Experiment 3: Accuracy Model for Words (blmm_exp3_acc) ---

exp3_eff_acc_agr_vs_dis <- blmm_exp3_acc %>% brms_effect_summary("Condition1")
exp3_eff_acc_gender <- blmm_exp3_acc %>% brms_effect_summary("StimulusType1")
exp3_eff_acc_agr_vs_dis_by_gender <- blmm_exp3_acc %>% brms_effect_summary("Condition1:StimulusType1")

```

Unlike in the first two experiments, we did not observe any evidence for an effect of the flanker condition in the mean of the Gaussian component of response time, $\mu_{RT}$ (`r exp3_eff_rt_mu_agr_vs_dis`) or the scale component, $\beta_{RT}$ (`r exp3_eff_rt_beta_agr_vs_dis`). There was also little evidence for effects of any of the manipulations on accuracy for word trials in Experiment 2 (`r exp3_eff_acc_agr_vs_dis`).

### Effects of grammatical gender

Just like in Experiment 1 (but unlike in Experiment 2), there was some evidence for a marginal effect of grammatical gender across flanker conditions in response times, with both the mean of the gaussian component shifted to the right for feminine compared to masculine target words ($\mu_{RT}$: `r exp3_eff_rt_mu_gender`) and the scale component for the exponential component being larger for feminine compared to masculine target words ($\beta_{RT}$: `r exp3_eff_rt_beta_gender`), indicating a longer right tail of the distribution. This is the opposite pattern to the one observed in Experiment 1, where responses to feminine target words were slightly faster than responses to masculine target words. However, there was no evidence for a gender effect on accuracy (`r exp3_eff_acc_gender`).

### Interaction between flanker condition and grammatical gender

We did not observe meaningful interactions between grammatical gender of the target word and the flanker condition in response time or accuracy ($\mu_{RT}$: `r exp3_eff_rt_mu_agr_vs_dis_by_gender`; $\beta_{RT}$: `r exp3_eff_rt_beta_agr_vs_dis_by_gender`; $\theta_{accuracy}$: `r exp3_eff_acc_agr_vs_dis_by_gender`).

## Discussion
Unlike in the previous experiments, we did not observe an effect of the article flanker condition in Experiment 3. This suggests that, for certain tasks, joint processing of article and noun is either not mandatory, or the joint processing does not influence the response. It seems that a semantic judgment task is high-level enough that the grammatical gender of the target word is either not processed or does not interfere with the semantic decision. This is again, in general terms, in line with the model by @sá-leite2024. We did observe a slight facilitation in terms of response times for masculine words compared to feminine words, which, unlike the findings in Experiments 1 and 2, is in line with the “masculine as default grammatical gender” hypothesis. It is not completely clear why a semantic categorization would be easier for masculine words than for feminine words, while decisions about grammatical gender seemed to be easier for feminine words and feminine pseudowords seemed to be harder to reject. Perhaps what we see in Experiments 1 and 2 is the result of participants attempting to compensate for a bias towards masculine words, resulting in them processing masculine and apparent masculine words more carefully, while this was not necessary in a semantic categorization task. In any case, the main focus of our study is the effect of the gender agreement manipulation.  We will discuss the implications of this finding together with those from Experiments 1 and 2 in the following.
